---
title: "Hands-on_Ex15"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **15  Choropleth Mapping with R**

## **15.1 Overview**

Choropleth mapping is a technique for visualizing data across geographic areas—like countries, states, or census units—by using area patterns or graduated colors to represent different values. For example, a social scientist might create a choropleth map to show the spatial distribution of Singapore's aged population by using the Master Plan 2014 Subzone Boundary data. In this chapter, you'll learn how to create both functional and accurate choropleth maps with the R package **tmap**.

## **15.2 Getting Started**

This section introduces the setup for a hands-on exercise in R for map visualization. The key points are:

-   **Core Packages:**

    -   **tmap:** Used for creating choropleth maps and other map visualizations.

    -   **readr, tidyr, dplyr:** Part of the tidyverse, used for importing delimited text files, tidying, and wrangling data.

    -   **sf:** Used for handling geospatial data.

## **15.3 Importing Data into R**

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

The chapter describes two datasets used for creating the choropleth map:

-   Master Plan 2014 Subzone Boundary (Web):\
    This dataset is an ESRI shapefile available from data.gov.sg. It contains the geographic boundaries of Singapore's planning subzones, based on URA Master Plan 2014.

-   Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020:\
    This dataset is in CSV format and is available from the Department of Statistics, Singapore. Although it does not include coordinate values, the PA (Planning Area) and SZ (Subzone) fields can be used as unique identifiers to geocode this data and join it with the shapefile.

These two datasets are combined in R to produce a choropleth map that visually represents the spatial distribution of data across Singapore.

### **15.3.2 Importing Geospatial Data into R**

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz
```

### **15.3.3 Importing Attribute Data into R**

Next, we will import *respopagsex2011to2020.csv* file into RStudio and save the file into an R dataframe called *popagsex*.

The task will be performed by using *read_csv()* function of **readr**package as shown in the code chunk below.

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### **15.3.4 Data Preparation**

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,

-   AGED: age group 65 and above,

-   TOTAL: all age group, and

-   DEPENDENCY: the ratio between young and aged against economy active group

#### 15.3.4.1 Data wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

#### 15.3.4.2 Joining the attribute data and geospatial data

Before joining the data, convert the PA and SZ fields to uppercase to match the already uppercase geospatial fields (SUBZONE_N and PLN_AREA_N).

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, *left_join()* of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ*as the common identifier.

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

Thing to learn from the code chunk above:

-   *left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## **15.4 Choropleth Mapping Geospatial Data Using *tmap***

Two approaches can be used to prepare thematic map using *tmap*, they are:

-   Plotting a thematic map quickly by using *qtm()*.

-   Plotting highly customisable thematic map by using tmap elements.

### **15.4.1 Plotting a choropleth map quickly by using *qtm()***

The easiest and quickest to draw a choropleth map using **tmap** is using *qtm()*. It is concise and provides a good default visualisation in many cases.

The code chunk below will draw a cartographic standard choropleth map as shown below.

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

Things to learn from the code chunk above:

-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)

### **15.4.2 Creating a choropleth map by using *tmap*’s elements**

Despite its usefulness of drawing a choropleth map quickly and easily, the disadvantge of *qtm()* is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

In the following sub-section, we will share with you tmap functions that used to plot these elements.

#### 15.4.2.1 Drawing a base map

The basic building block of **tmap** is *tm_shape()* followed by one or more layer elemments such as *tm_fill()* and *tm_polygons()*.

In the code chunk below, *tm_shape()* is used to define the input data (i.e *mpsz_pop2020*) and *tm_polygons()* is used to draw the planning subzone polygons

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 15.4.2.2 Drawing a choropleth map using *tm_polygons()*

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, we just need to assign the target variable such as *Dependency* to *tm_polygons()*.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

-   **Default Binning:** The tm_polygons() function uses a "pretty" interval binning method by default. (More details in sub-section 4.3.)

-   **Default Color Scheme:** It applies the YlOrRd color scheme from ColorBrewer. (See sub-section 4.4 for more information.)

-   **Handling Missing Values:** Missing values are shaded in grey by default.

#### 15.4.2.3 Drawing a choropleth map using *tm_fill()* and \*tm_border()\*\*

**tm_polygons()** is essentially a wrapper that combines tm_fill() and tm_border(). The tm_fill() function applies the default color scheme to fill the polygons, while tm_border() adds the shapefile's borders to the map. The example code demonstrates drawing a choropleth map using only tm_fill(), which means the map will display colored polygons without the border lines.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

Notice that the planning subzones are shared according to the respective dependecy values

To add the boundary of the planning subzones, tm_borders will be used as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1, fill_alpha = 1) 
```

### **15.4.3 Data classification methods of tmap**

Most choropleth maps employ some methods of data classification. The point of classification is to take a large number of observations and group them into data ranges or classes.

**tmap** provides a total ten data classification methods, namely: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.

To define a data classification method, the *style* argument of *tm_fill()*or *tm_polygons()* will be used.

#### 15.4.3.1 Plotting choropleth maps with built-in classification methods

The code chunk below shows a quantile data classification that used 5 classes.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

In the code chunk below, *equal* data classification method is used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

Notice that the distribution of quantile data classification method are more evenly distributed then equal data classification method.\

::: callout-warning
## Warning: Maps Lie!

DIY: Using what you had learned, prepare choropleth maps by using different classification methods supported by tmap and compare their differences.

DIY: Preparing choropleth maps by using similar classification method but with different numbers of classes (i.e. 2, 6, 10, 20). Compare the output maps, what observation can you draw?
:::

#### 15.4.3.2 Plotting choropleth map with custome break

Built-in styles in tmap automatically calculate category breaks, but you can override these by explicitly setting the breakpoints using the breaks argument in tm_fill(). Note that to achieve n categories, you must specify n+1 breakpoints (in increasing order, including both the minimum and maximum values). It is also recommended to review descriptive statistics of the variable (for example, the DEPENDENCY field) before setting these custom breakpoints.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

With reference to the results above, we set break point at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which we set at 0 and 100. Our *breaks* vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00)

Now, we will plot the choropleth map by using the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

### **15.4.4 Colour Scheme**

**tmap** supports colour ramps either defined by the user or a set of predefined colour ramps from the **RColorBrewer** package.

#### 15.4.4.1 Using ColourBrewer palette

To change the colour, we assign the preferred colour to *palette*argument of *tm_fill()* as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

Notice that the choropleth map is shaded in green.

To reverse the colour shading, add a “-” prefix.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

Notice that the colour scheme has been reversed.

### **15.4.5 Map Layouts**

Map layouts refer to the overall composition of the map, combining all its elements—such as the mapped data, title, scale bar, compass, margins, and aspect ratio—into a cohesive design. Additionally, the colour settings and data classification methods (i.e., the palette and breakpoints) play a crucial role in defining the map’s visual appearance.

#### 15.4.5.1 Map Legend

In **tmap**, several *legend* options are provided to change the placement, format and appearance of the legend.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

#### 15.4.5.2 Map style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.

The code chunk below shows the *classic* style is used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### 15.4.5.3 Cartographic Furniture

Beside map style, **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

In the code chunk below, *tm_compass()*, *tm_scale_bar()* and *tm_grid()*are used to add compass, scale bar and grid lines onto the choropleth map.

```{r}
#| fig.width: 12
#| fig.height: 7
#| echo: true
#| message: false
#| warning: false

tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY", 
          fill.scale = tm_scale_intervals(style = "quantile", 
                                          values = "brewer.blues"),
          fill.legend = tm_legend(title = "No. of persons")) +  
  tm_title("Distribution of Dependency Ratio \nby planning subzone") +  
  tm_layout(
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45,
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type = "8star", size = 2) +
  tm_scalebar(position = c("right", "bottom")) +  
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authority (URA)\n and Population data from Department of Statistics DOS",
             position = c("left", "bottom"))
```

To reset the default style, refer to the code chunk below.

```{r}
tmap_style("white")
```

### **15.4.6 Drawing Small Multiple Choropleth Maps**

**Small multiple maps**, also referred to as **facet maps**, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In **tmap**, small multiple maps can be plotted in three ways:

-   by assigning multiple values to at least one of the asthetic arguments,

-   by defining a group-by variable in *tm_facets()*, and

-   by creating multiple stand-alone maps with *tmap_arrange()*.

#### 15.4.6.1 By assigning multiple values to at least one of the aesthetic arguments

In this example, small multiple choropleth maps are created by defining ***ncols*** in **tm_fill()**

```{r}
#| fig.width: 12
#| fig.height: 7
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments

```{r}
#| fig.width: 12
#| fig.height: 7
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### 15.4.6.2 By defining a group-by variable in *tm_facets()*

In this example, multiple small choropleth maps are created by using **tm_facets()**.

```{r}
#| fig.width: 7
#| fig.height: 5
#| message: false
#| warning: false
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY", 
          fill.scale = tm_scale_intervals(style = "quantile"),  
          values = "brewer.blues",  
          thres.poly = 0) + 
  tm_facets(by = "REGION_N", 
            free.coords = FALSE,  
            drop.units = TRUE) +  
  tm_layout(
            legend.show = FALSE,
            title.position = "center",  
            title.size = 1.5,  
            frame = TRUE) +
  tm_borders(fill_alpha = 1) +  
  tm_compass(type = "8star", size = 2) +  
  tm_scalebar(position = c("right", "bottom")) +  
  tm_grid(lwd = 0.1, alpha = 0.2) +  
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authority (URA)\n and Population data from Department of Statistics DOS",
             position = c("left", "bottom"))

```

#### 15.4.6.3 By creating multiple stand-alone maps with *tmap_arrange()*

In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with **tmap_arrange()**.

```{r}
#| fig.width: 12
#| fig.height: 7
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### **15.4.7 Mappping Spatial Object Meeting a Selection Criterion**

Instead of creating small multiple choropleth map, you can also use selection funtion to map spatial objects meeting the selection criterion.

```{r}
#| fig.width: 12
#| fig.height: 7
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

## **15.5 Reference**

### **15.5.1 All about tmap package**

-   [tmap: Thematic Maps in R](https://www.jstatsoft.org/article/view/v084i06)

-   [tmap](https://cran.r-project.org/web/packages/tmap/index.html)

-   [tmap: get started!](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)

-   [tmap: changes in version 2.0](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-changes-v2.html)

-   [tmap: creating thematic maps in a flexible way (useR!2015)](http://von-tijn.nl/tijn/research/presentations/tmap_user2015.pdf)

-   [Exploring and presenting maps with tmap (useR!2017)](http://von-tijn.nl/tijn/research/presentations/tmap_user2017.pdf)

### **15.5.2 Geospatial data wrangling**

-   [sf: Simple Features for R](https://cran.r-project.org/web/packages/sf/index.html)

-   [Simple Features for R: StandardizedSupport for Spatial Vector Data](https://journal.r-project.org/archive/2018/RJ-2018-009/RJ-2018-009.pdf)

-   [Reading, Writing and Converting Simple Features](https://cran.r-project.org/web/packages/sf/vignettes/sf2.html)

### **15.5.3 Data wrangling**

-   [dplyr](https://dplyr.tidyverse.org/)

-   [Tidy data](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html)

-   [tidyr: Easily Tidy Data with ‘spread()’ and ‘gather()’ Functions](https://cran.r-project.org/web/packages/tidyr/tidyr.pdf)
