---
title: "Take-home Exercise 1"
author: "Jeounghyun Kim"
date: "February 1, 2025"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **Heart Attack Risks: A Data-Driven Analysis**

# **1. Overview**

## 1.1 Introduction

According to the World Health Organization (WHO), cardiovascular diseases (CVDs) are the leading cause of death globally, accounting for approximately 17.9 million deaths per year. Among them, heart attacks are particularly critical, emphasizing the need for early prevention and risk analysis.

This dataset examines heart attack incidents in Japan, comparing youth and adult populations. It includes medical indicators (cholesterol, blood pressure, heart rate), lifestyle factors (exercise, diet), and demographics, offering insights into age-specific cardiovascular risks.

By leveraging this dataset, researchers and policymakers can enhance prevention strategies and compare findings with global initiatives like WHO’s *Global Hearts* and the CDC’s *Million Hearts* program. It also supports statistical modeling and predictive analytics to inform data-driven public health policies and early interventions.

## 1.2 The Data

The data for this analysis comes from the **Heart Attack in Japan: Youth vs. Adult** dataset, retrieved from the [Kaggle database](https://www.kaggle.com/datasets/ashaychoudhary/heart-attack-in-japan-youth-vs-adult/data). The dataset provides detailed information on heart attack incidents among youth and adults, focusing on medical indicators, lifestyle factors, and demographic characteristics.

**Dataset Components**

The dataset includes various features, such as:

-   **Demographic Information:** Age, gender, and residential region (Urban/Rural).

-   **Medical History:** Presence of pre-existing conditions such as **diabetes** and **hypertension**.

-   **Lifestyle Factors:**

    -   **Smoking History** and **Alcohol Consumption** levels.

    -   **Physical Activity** categorized as High, Moderate, or Low.

    -   **Diet Quality** classified as Good, Average, or Poor.

-   **Medical Indicators:**

    -   **Cholesterol Levels** (numerical values).

    -   **Resting Blood Pressure**

-   Additionally, the dataset contains extra computed columns that are not directly relevant to this analysis. These columns will be removed during data preprocessing to ensure cleaner and more interpretable results.

# **2. Getting started**

## 2.1 Load Packages

We load the following R packages using the **`pacman::p_load()`** function to streamline the workflow:

-   **tidyverse**: A core collection of R packages designed for data science, including data manipulation (`dplyr`), visualization (`ggplot2`), and data tidying (`tidyr`).

<!-- -->

-   **ggiraph**: Provides interactive `ggplot2` visualizations.

-   **plotly**: Enables interactive plots for data exploration.

-   **patchwork**: Facilitates combining multiple `ggplot2` visualizations into a cohesive layout.

-   **DT**: Provides an interactive table interface for data visualization.

-   **GGally**: Extends `ggplot2` for correlation and pairwise visualizations.

-   **zoo**: Provides functions for time series data manipulation and rolling window calculations.

-   **ggcorrplot**: Specialized in correlation matrix visualization using `ggplot2`.

-   **ggstatsplot**: Provides statistical visualizations, integrating hypothesis testing within `ggplot2`.

-   **scales**: Enhances `ggplot2` with functions for scaling and formatting axes, labels, and colors.

```{r}
pacman::p_load(ggiraph, plotly, patchwork, DT, tidyverse, 
               GGally, zoo, ggcorrplot, ggstatsplot, scales)

```

## 2.2 Import data

```{r}
df <- read_csv("japan_heart_attack_dataset.csv")
```

## 2.3 Data Overview & pre-processing

This output from `glimpse(df)` provides a summary of the dataset's structure, offering an overview of the number of rows, columns, and the types of variables included.

```{r}
glimpse(df)
```

### 2.3.1 Check missing values

`colSums(is.na(df))` calculates the **number of missing values** in each column of the dataset.

```{r}

colSums(is.na(df))
```

### 2.3.2 Data Cleaning and Encoding of Categorical Variables

This preprocessing step transforms categorical variables into numeric representations and removes unnecessary columns to prepare the dataset for analysis.

```{r}
df <- df %>%
  mutate(
    smoking_history = ifelse(Smoking_History == "Yes", 1, 0),
    diabetes_history = ifelse(Diabetes_History == "Yes", 1, 0),
    hypertension_history = ifelse(Hypertension_History == "Yes", 1, 0),
    gender = ifelse(Gender == "Male", 1, 0),
    heart_attack_occurrence = ifelse(Heart_Attack_Occurrence == "Yes", 1, 0),
    region = ifelse(Region == "Urban", 1, 0),
    family_history = ifelse(Family_History == "Yes", 1, 0), 
    
    physical_activity = case_when(
      Physical_Activity == "Low" ~ 0,
      Physical_Activity == "Moderate" ~ 1,
      Physical_Activity == "High" ~ 2
    ),
    
    diet_quality = case_when(
      Diet_Quality == "Poor" ~ 0,
      Diet_Quality == "Average" ~ 1,
      Diet_Quality == "Good" ~ 2
    ),
    
    alcohol_consumption = case_when(
      Alcohol_Consumption == "None" ~ 0, 
      Alcohol_Consumption == "Low" ~ 1,   
      Alcohol_Consumption == "Moderate" ~ 2,
      Alcohol_Consumption == "High" ~ 3,
    )
  ) %>%
  select(-Smoking_History, -Diabetes_History, -Hypertension_History, 
         -Gender, -Heart_Attack_Occurrence, -Region, -Family_History,
         -Physical_Activity, -Diet_Quality, -Alcohol_Consumption,
         -starts_with("Extra_Column_")) 
```

**- Encoding Categorical Variables**

-   Binary Encoding (0/1):

    -   `smoking_history`, `diabetes_history`, `hypertension_history`, `gender`, `heart_attack_occurrence`, `region`, `family_history` are converted into binary numeric values (`1 = Yes/Male/Urban`, `0 = No/Female/Rural`).

-   Ordinal Encoding:

    -   `physical_activity`: Encoded as `0 = Low`, `1 = Moderate`, `2 = High`.

    -   `diet_quality`: Encoded as `0 = Poor`, `1 = Average`, `2 = Good`.

    -   `alcohol_consumption`: Encoded as `0 = None`, `1 = Low`, `2 = Moderate`, `3 = High`.

**- Removing Unnecessary Columns**

-   The original categorical columns (`Smoking_History`, `Diabetes_History`, etc.) are removed after encoding.

-   Extra computed columns (`Extra_Column_1` to `Extra_Column_15`) are removed as they are not relevant to the analysis.

**- Purpose of This Preprocessing**

-   Makes data suitable for statistical modeling and machine learning, which typically require numeric inputs.

-   Reduces dimensionality by eliminating redundant columns.

-   Ensures consistency by encoding categorical variables in a structured way.

```{r}
ls(df)
```

`ls(df)` **lists all column names in the processed dataset**, allowing verification of the cleaned and transformed variables after preprocessing.

# **3. Exploratory Data Analysis**

## 3.1 Heart Attack Occurrence Rate by Age (5-Year Rolling Average)

#### **Graph Description**

The graph visualizes the heart attack occurrence rate across different ages, with a 5-year rolling average applied to smooth fluctuations. The red line represents the raw occurrence rate, while the blue line highlights the trend. The rate generally increases with age, but interestingly, there are notable variations across different age groups.

::: panel-tabset
#### The Plot

```{r}
#| echo: false
#library(dplyr)
#library(ggplot2)
#library(zoo)

age_heart_attack_rate <- df %>%
  group_by(Age) %>%
  summarize(heart_attack_rate = mean(heart_attack_occurrence, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Age = as.numeric(Age))  

age_heart_attack_rate <- age_heart_attack_rate %>%
  arrange(Age) %>%
  mutate(rolling_avg = rollmean(heart_attack_rate, k = 5, fill = mean(heart_attack_rate, na.rm = TRUE), align = "right", na.rm = TRUE))

ggplot(na.omit(age_heart_attack_rate), aes(x = Age)) +
  geom_line(aes(y = heart_attack_rate), color = "red", size = 1, alpha = 0.5) +  
  geom_line(aes(y = rolling_avg), color = "blue", size = 1.2) + 
  geom_point(aes(y = rolling_avg), color = "blue", size = 2) +  
  labs(title = "Heart Attack Occurrence Rate by Age (5-Year Rolling Avg)",
       x = "Age",
       y = "Heart Attack Occurrence (%)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed")
  )
```

#### The code chunk

```{r, echo=TRUE, eval=FALSE}
#library(dplyr)
#library(ggplot2)
#library(zoo)

age_heart_attack_rate <- df %>%
  group_by(Age) %>%
  summarize(heart_attack_rate = mean(heart_attack_occurrence, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Age = as.numeric(Age))  

age_heart_attack_rate <- age_heart_attack_rate %>%
  arrange(Age) %>%
  mutate(rolling_avg = rollmean(heart_attack_rate, k = 5, fill = mean(heart_attack_rate, na.rm = TRUE), align = "right", na.rm = TRUE))

ggplot(na.omit(age_heart_attack_rate), aes(x = Age)) +
  geom_line(aes(y = heart_attack_rate), color = "red", size = 1, alpha = 0.5) +  
  geom_line(aes(y = rolling_avg), color = "blue", size = 1.2) + 
  geom_point(aes(y = rolling_avg), color = "blue", size = 2) +  
  labs(title = "Heart Attack Occurrence Rate by Age (5-Year Rolling Avg)",
       x = "Age",
       y = "Heart Attack Occurrence (%)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed")
  )
```
:::

#### **Observation**

An unexpectedly high occurrence is observed in individuals in their 30s and 40s, suggesting that younger adults may also be at significant risk. This contradicts the common perception that heart attacks predominantly affect older populations. Potential contributing factors include increased stress levels, smoking, and undiagnosed pre-existing conditions. The temporary decline in heart attack rates in the late 50s may reflect the impact of medical interventions or lifestyle modifications, while the sharp increase beyond the late 60s aligns with conventional risk factors such as hypertension, diabetes, and cardiovascular aging.

## 3.2 Heart Attack and Risk Factors by Age Group (10-Year Intervals)

#### **Graph Description**

This graph illustrates the heart attack rate along with key risk factors—smoking, hypertension, and diabetes—across different age groups in 10-year intervals. Each colored line represents a different variable, highlighting how these factors change with age.

::: panel-tabset
#### The Plot

```{r}
#| echo: false

df %>%
  mutate(Age_Group = cut(Age, 
                         breaks = c(0, 20, 30, 40, 50, 60, 70, 80, Inf),  
                         include.lowest = TRUE, 
                         right = FALSE, 
                         labels = c("<20", "20s", "30s", "40s", "50s", "60s", "70s", "80+"))) %>%
  mutate(Age_Group = factor(Age_Group, levels = c("<20", "20s", "30s", "40s", "50s", "60s", "70s", "80+"))) %>%
  group_by(Age_Group) %>%
  summarise(
    heart_attack_rate = mean(heart_attack_occurrence, na.rm = TRUE) * 100,
    smoking_rate = mean(smoking_history, na.rm = TRUE) * 100,  
    hypertension_rate = mean(hypertension_history, na.rm = TRUE) * 100,  
    diabetes_rate = mean(diabetes_history, na.rm = TRUE) * 100  
  ) %>%
  pivot_longer(cols = c(heart_attack_rate, smoking_rate, hypertension_rate, diabetes_rate),
               names_to = "variable", values_to = "rate") %>%
  ggplot(aes(x = Age_Group, y = rate, color = variable, group = variable)) +
  geom_line(size = 1.2) +  
  geom_point(size = 3) +  
  labs(title = "Heart Attack and Risk Factors by Age Group (10-year intervals)",
       x = "Age Group",
       y = "Rate (%)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(size = 12), 
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed")
  )
```

#### The code chunk

```{r, echo=TRUE, eval=FALSE}

df %>%
  mutate(Age_Group = cut(Age, 
                         breaks = c(0, 20, 30, 40, 50, 60, 70, 80, Inf),  
                         include.lowest = TRUE, 
                         right = FALSE, 
                         labels = c("<20", "20s", "30s", "40s", "50s", "60s", "70s", "80+"))) %>%
  mutate(Age_Group = factor(Age_Group, levels = c("<20", "20s", "30s", "40s", "50s", "60s", "70s", "80+"))) %>%
  group_by(Age_Group) %>%
  summarise(
    heart_attack_rate = mean(heart_attack_occurrence, na.rm = TRUE) * 100,
    smoking_rate = mean(smoking_history, na.rm = TRUE) * 100,  
    hypertension_rate = mean(hypertension_history, na.rm = TRUE) * 100,  
    diabetes_rate = mean(diabetes_history, na.rm = TRUE) * 100  
  ) %>%
  pivot_longer(cols = c(heart_attack_rate, smoking_rate, hypertension_rate, diabetes_rate),
               names_to = "variable", values_to = "rate") %>%
  ggplot(aes(x = Age_Group, y = rate, color = variable, group = variable)) +
  geom_line(size = 1.2) +  
  geom_point(size = 3) +  
  labs(title = "Heart Attack and Risk Factors by Age Group (10-year intervals)",
       x = "Age Group",
       y = "Rate (%)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(size = 12), 
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_line(color = "gray", linetype = "dashed")
  )
```
:::

**Observation**

-   Hypertension and diabetes rates tend to increase steadily until the 50s, followed by a slight decline after 60, which may be influenced by increased medical intervention and lifestyle modifications. Smoking rates peak in the 20s and 30s before declining sharply after 60, likely due to health concerns or cessation efforts.

-   Although smoking rates decrease in older age groups, earlier exposure to smoking, along with increasing hypertension and diabetes prevalence, may be associated with cardiovascular risks in individuals in their 30s–50s. Conversely, in older adults (60+), chronic conditions such as diabetes and hypertension appear to remain relevant factors for heart attack risk, even as smoking rates decline.

## 3.3 BMI Distribution by Age Group (30-40s vs. 60+)

#### **Graph Description**

This violin plot with a boxplot overlay compares the BMI distributions of individuals aged 30-40s and 60+. It provides a visual representation of the median values, spread, and density for each group.

::: panel-tabset
#### The Plot

```{r}
#| echo: false
df %>%
  filter(Age >= 30 & Age <= 40 | Age >= 60) %>%
  mutate(
    age_group = case_when(
      Age >= 30 & Age <= 40 ~ "30-40s",
      Age >= 60 ~ "60+"
    ),
    heart_attack_label = ifelse(heart_attack_occurrence == 1, "Yes", "No")
  ) %>%
  ggbetweenstats(
    x = age_group, 
    y = BMI, 
    grouping.var = heart_attack_label,  
    type = "np",                        # Mann-Whitney U test
    pairwise.comparisons = TRUE,        
    messages = FALSE
  )

```

#### The code chunk

```{r, echo=TRUE, eval=FALSE}
df %>%
  filter(Age >= 30 & Age <= 40 | Age >= 60) %>%
  mutate(
    age_group = case_when(
      Age >= 30 & Age <= 40 ~ "30-40s",
      Age >= 60 ~ "60+"
    ),
    heart_attack_label = ifelse(heart_attack_occurrence == 1, "Yes", "No")
  ) %>%
  ggbetweenstats(
    x = age_group, 
    y = BMI, 
    grouping.var = heart_attack_label,  
    type = "np",                       # Mann-Whitney U test
    pairwise.comparisons = TRUE,       
    messages = FALSE
  )

```
:::

#### **Observation**

-   The median BMI is 24.90 for individuals aged 30-40 and 24.97 for those aged 60+, indicating no statistically significant difference. A Mann-Whitney U test confirms this (W = 2.70×10⁷, p = 0.99), with a rank biserial correlation of nearly zero (-1.07e-04) and a confidence interval of \[-0.02, 0.02\], suggesting that BMI does not serve as a key differentiator for heart attack risk between these age groups.
-   Despite the common perception that BMI influences heart attack risk, this analysis suggests that other factors, such as cholesterol levels, physical activity, and stress, may play a more significant role in determining cardiovascular outcomes.

## 3.4 Heart Attack Rate by Alcohol Consumption Level

#### **Graph Description**

This bar chart illustrates the heart attack rates across different levels of alcohol consumption: None, Low, Moderate, and High. The data suggests a non-linear trend, with moderate drinkers showing the lowest heart attack rate, while both non-drinkers and heavy drinkers exhibit slightly higher rates.

::: panel-tabset
#### The Plot

```{r}
#| echo: false
#library(dplyr)
#library(ggplot2)

heart_attack_by_alcohol <- df %>%
  group_by(alcohol_consumption) %>%
  summarize(heart_attack_rate = mean(heart_attack_occurrence, na.rm = TRUE)) %>%
  ungroup()


heart_attack_by_alcohol <- heart_attack_by_alcohol %>%
  mutate(
    alcohol_label = factor(alcohol_consumption, 
                           levels = c(0, 1, 2, 3), 
                           labels = c("None", "Low", "Moderate", "High"))
  )

ggplot(heart_attack_by_alcohol, aes(x = alcohol_label, y = heart_attack_rate, fill = alcohol_label)) +
  geom_bar(stat = "identity", width = 0.6, color = "black", size = 0.5) + 
  scale_fill_manual(values = c("None" = "#A93226", "Low" = "#E67E22", "Moderate" = "#2ECC71", "High" = "#3498DB")) + # 색상 강조
  labs(
    title = "Heart Attack Rate by Alcohol Consumption Level",
    x = "Alcohol Consumption Level",
    y = "Heart Attack Rate (%)",
    fill = "Alcohol Level"
  ) +
  theme_minimal(base_size = 14) + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
    axis.title.x = element_text(size = 14, face = "bold"), 
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12), 
    legend.title = element_text(size = 14, face = "bold"), 
    legend.text = element_text(size = 12), 
    panel.grid.major = element_line(color = "gray90"), 
    panel.grid.minor = element_blank()
  )
```

#### The code chunk

```{r, echo=TRUE, eval=FALSE}
library(dplyr)
library(ggplot2)

heart_attack_by_alcohol <- df %>%
  group_by(alcohol_consumption) %>%
  summarize(heart_attack_rate = mean(heart_attack_occurrence, na.rm = TRUE)) %>%
  ungroup()


heart_attack_by_alcohol <- heart_attack_by_alcohol %>%
  mutate(
    alcohol_label = factor(alcohol_consumption, 
                           levels = c(0, 1, 2, 3), 
                           labels = c("None", "Low", "Moderate", "High"))
  )

ggplot(heart_attack_by_alcohol, aes(x = alcohol_label, y = heart_attack_rate, fill = alcohol_label)) +
  geom_bar(stat = "identity", width = 0.6, color = "black", size = 0.5) + 
  scale_fill_manual(values = c("None" = "#A93226", "Low" = "#E67E22", "Moderate" = "#2ECC71", "High" = "#3498DB")) + # 색상 강조
  labs(
    title = "Heart Attack Rate by Alcohol Consumption Level",
    x = "Alcohol Consumption Level",
    y = "Heart Attack Rate (%)",
    fill = "Alcohol Level"
  ) +
  theme_minimal(base_size = 14) + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
    axis.title.x = element_text(size = 14, face = "bold"), 
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12), 
    legend.title = element_text(size = 14, face = "bold"), 
    legend.text = element_text(size = 12), 
    panel.grid.major = element_line(color = "gray90"), 
    panel.grid.minor = element_blank()
  )
```
:::

#### **Observation**

-   The findings suggest that moderate alcohol consumption may be linked to a lower risk of heart attack, potentially due to improved circulation or reduced stress. However, this does not imply that drinking is beneficial. Other contributing factors, such as lifestyle, diet, and pre-existing health conditions, could influence these results.

<!-- -->

-   Further investigation is required to determine whether alcohol consumption has a direct causal effect on heart attack risk or if other confounding factors play a more significant role.

## 3.5 Standardized Health Factors by Heart Attack Occurrence (by Age Group)

#### **Graph Description**

This bar chart displays the standardized Z-scores of key health factors for individuals with and without heart attacks, categorized into two age groups: 30-40s and 60+. Red bars represent those who have experienced a heart attack, while blue bars represent those who have not.

::: panel-tabset
#### The Plot

```{r}
#| echo: false
df_scaled <- df %>%
  mutate(age_group = ifelse(Age >= 60, "60+", "30-40s")) %>%
  group_by(age_group) %>%
  mutate(
    diet_quality_scaled = scale(diet_quality),
    physical_activity_scaled = scale(physical_activity),
    BMI_scaled = scale(BMI),
    stress_scaled = scale(Stress_Levels),
    cholesterol_scaled = scale(Cholesterol_Level),
    systolic_BP_scaled = scale(Systolic_BP),
    diastolic_BP_scaled = scale(Diastolic_BP)
  ) %>%
  ungroup()

health_factors_by_age <- df_scaled %>%
  group_by(age_group, heart_attack_occurrence) %>%
  summarise(
    diet_quality_avg = mean(diet_quality_scaled),
    physical_activity_avg = mean(physical_activity_scaled),
    BMI_avg = mean(BMI_scaled),
    stress_avg = mean(stress_scaled),
    cholesterol_avg = mean(cholesterol_scaled),
    systolic_BP_avg = mean(systolic_BP_scaled),
    diastolic_BP_avg = mean(diastolic_BP_scaled)
  ) %>%
  pivot_longer(cols = -c(age_group, heart_attack_occurrence), 
               names_to = "variable", values_to = "value")

ggplot(health_factors_by_age, aes(x = variable, y = value, fill = as.factor(heart_attack_occurrence))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~age_group) +  
  labs(title = "Standardized Health Factors by Heart Attack Occurrence (by Age Group)",
       x = "Health Factor",
       y = "Standardized Average Value (Z-score)",
       fill = "Heart Attack (0: No, 1: Yes)") +
  scale_fill_manual(values = c("0" = "#3498DB", "1" = "#E74C3C")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),  
    axis.title.x = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    legend.title = element_text(face = "bold"),
    strip.text = element_text(size = 12, face = "bold")  
  )
```

#### The code chunk

```{r, echo=TRUE, eval=FALSE}
df_scaled <- df %>%
  mutate(age_group = ifelse(Age >= 60, "60+", "30-40s")) %>%
  group_by(age_group) %>%
  mutate(
    diet_quality_scaled = scale(diet_quality),
    physical_activity_scaled = scale(physical_activity),
    BMI_scaled = scale(BMI),
    stress_scaled = scale(Stress_Levels),
    cholesterol_scaled = scale(Cholesterol_Level),
    systolic_BP_scaled = scale(Systolic_BP),
    diastolic_BP_scaled = scale(Diastolic_BP)
  ) %>%
  ungroup()

health_factors_by_age <- df_scaled %>%
  group_by(age_group, heart_attack_occurrence) %>%
  summarise(
    diet_quality_avg = mean(diet_quality_scaled),
    physical_activity_avg = mean(physical_activity_scaled),
    BMI_avg = mean(BMI_scaled),
    stress_avg = mean(stress_scaled),
    cholesterol_avg = mean(cholesterol_scaled),
    systolic_BP_avg = mean(systolic_BP_scaled),
    diastolic_BP_avg = mean(diastolic_BP_scaled)
  ) %>%
  pivot_longer(cols = -c(age_group, heart_attack_occurrence), 
               names_to = "variable", values_to = "value")

ggplot(health_factors_by_age, aes(x = variable, y = value, fill = as.factor(heart_attack_occurrence))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~age_group) +  
  labs(title = "Standardized Health Factors by Heart Attack Occurrence (by Age Group)",
       x = "Health Factor",
       y = "Standardized Average Value (Z-score)",
       fill = "Heart Attack (0: No, 1: Yes)") +
  scale_fill_manual(values = c("0" = "#3498DB", "1" = "#E74C3C")) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),  
    axis.title.x = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    legend.title = element_text(face = "bold"),
    strip.text = element_text(size = 12, face = "bold")  
  )
```
:::

#### **Observation**

-   30-40s: Higher stress levels and lower physical activity appear to be more common among individuals who experienced a heart attack. Elevated cholesterol levels and poor diet quality are also observed to some extent, suggesting that lifestyle-related factors may be associated with cardiovascular risk in younger adults.

-   60+: Individuals who had a heart attack tend to show higher systolic blood pressure compared to those who did not. Unlike younger individuals, where lifestyle factors seem more relevant, chronic conditions such as hypertension and cholesterol levels may play a larger role in cardiovascular risks for older adults.

-   For younger individuals (30-40s), stress levels and physical activity may have some association with heart attack occurrence, while for older adults (60+), managing chronic conditions such as high blood pressure and cholesterol levels could be more relevant in heart attack prevention. However, these observations do not indicate direct causality, and further analysis would be needed to confirm stronger relationships.

# **4. Conclusion**

This analysis explored the relationship between heart attack occurrence across different age groups and various health factors, including blood pressure, cholesterol levels, stress, and lifestyle habits.

-   For individuals in their 30s and 40s, higher stress levels, lower physical activity, and elevated cholesterol levels were associated with a higher likelihood of heart attacks.

-   For those aged 60 and above, high blood pressure and elevated cholesterol levels were more prominent health indicators linked to heart attack occurrence.

Additionally, the group with moderate alcohol consumption showed the lowest heart attack rates. However, this does not indicate a direct causal relationship but rather suggests potential associations influenced by lifestyle habits and other health factors.

Based on these findings, stress management and lifestyle improvements may be crucial for younger individuals (30s–40s), while chronic disease prevention and management should be prioritized for those aged 60 and above. However, further research is necessary to establish more definitive causal relationships.

## 4.1 Future research Agenda

-   Further research should incorporate longitudinal studies to better understand the causal relationship between stress, physical activity, and heart disease in younger populations. Examining how lifestyle changes over time impact cardiovascular health will provide a deeper understanding of early prevention strategies.

-   Additionally, research should explore the effectiveness of policy interventions aimed at stress reduction and physical activity promotion among working-age adults. Investigating whether workplace wellness programs or public health initiatives can mitigate heart attack risks will be valuable.

-   For older populations, future studies should focus on optimizing healthcare strategies for the early detection and management of hypertension and cholesterol levels. Understanding the effectiveness of targeted medical interventions and personalized treatment approaches will be essential in preventing heart attacks in aging populations.

# **5. Key References**

-   [R for Visual Analytics](https://r4va.netlify.app/).

-   [R for Data Science](https://r4ds.hadley.nz/).

-   [Kaggle Japan Heart attack database](https://www.kaggle.com/datasets/ashaychoudhary/heart-attack-in-japan-youth-vs-adult/data)
