---
title: "Hands-on_Ex19"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## **19.1 Overview**

By the end of this hands-on exercise, you will be able to:

-   create bullet chart by using ggplot2,

-   create sparklines by using ggplot2 ,

-   build industry standard dashboard by using R Shiny.

## **19.2 Getting started**

For the purpose of this hands-on exercise, the following R packages will be used.

```{r}
#| message: false
#| warning: false
pacman::p_load(lubridate, ggthemes, reactable,
reactablefmtr, gt, gtExtras, tidyverse,RODBC)
```

The following R packages will be used for this exercise:

-   **tidyverse**: A collection of R packages for data science tasks such as importing, tidying, wrangling, and visualizing data (includes `readr`, `tidyr`, `dplyr`, `ggplot2`, and more).

-   **lubridate**: Simplifies working with dates and times.

-   **ggthemes**: Adds extra themes to `ggplot2` beyond the default options.

-   **gtExtras**: Provides helper functions to create visually appealing tables using the `gt` package.

-   **reactable**: Enables the creation of interactive data tables in R, based on the React Table library.

-   **reactablefmtr**: Enhances and customizes `reactable` tables easily with flexible formatting functions.

## **19.3 Importing Microsoft Access database**

### **19.3.1 The data set**

For the purpose of this study, a personal database in Microsoft Access mdb format called *Coffee Chain* will be used.

### **19.3.2 Importing database into R**

In the code chunk below, `odbcConnectAccess()` of [**RODBC**](https://cran.r-project.org/web/packages/RODBC/index.html) package is used used to import a database query table into R.

Note: Before running the code chunk, you need to change the R system to 32bit version. This is because the `odbcConnectAccess()` is based on 32bit and not 64bit

### **19.3.3 Data Preparation**

The code chunk below is used to import *CoffeeChain.rds* into R.

```{r}
coffeechain <- read_rds("data/CoffeeChain.rds")
```

Note: This step is optional if *coffeechain* is already available in R.

The code chunk below is used to aggregate Sales and Budgeted Sales at the Product level.

```{r}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

### **19.3.4 Bullet chart in ggplot2**

The code chunk below is used to plot the bullet charts using ggplot2 functions.

```{r}
#| echo: true
#| message: false
#| warning: false
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

## **19.4 Plotting sparklines using ggplot2**

In this section, you will learn how to plot sparklines by using ggplot2.

### **19.4.1 Preparing the data**

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

The code chunk below is used to compute the minimum, maximum and end othe the month sales.

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

The code chunk below is used to compute the 25 and 75 quantiles.

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)
```

### **19.4.2 sparklines in ggplot2**

The code chunk used.

```{r}
#| echo: true
#| message: false
#| warning: false
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

## **19.5 Static Information Dashboard Design: gt and gtExtras methods**

In this section, you’ll learn how to create **static information dashboards** using the **`gt`** and **`gtExtras`** packages.\
Before starting, it is **highly recommended** to visit the official webpages of both packages and review their materials.\
You **don’t need to memorize** everything—just get a general understanding of their **purpose and available functions**.

### **19.5.1 Plotting a simple bullet chart**

In this section, you will learn how to prepare a bullet chart report by using functions of gt and gtExtras packages.

```{r}
#| echo: true
#| message: false
#| warning: false
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## **19.6 sparklines: gtExtras method**

Before we can prepare the sales report by product by using gtExtras functions, code chunk below will be used to prepare the data.

```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()
```

It is important to note that one of the requirement of gtExtras functions is that almost exclusively they require you to pass data.frame with list columns. In view of this, code chunk below will be used to convert the report data.frame into list columns.

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

### **19.6.1 Plotting Coffechain Sales report**

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)
```

### **19.6.2 Adding statistics**

First, calculate summary statistics by using the code chunk below.

```{r}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

### **19.6.3 Combining the data.frame**

Next, use the code chunk below to add the statistics on the table.

```{r}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

```{r}
sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )
```

```{r}
sales_data = left_join(sales, spark)
```

### **19.6.4 Plotting the updated data.table**

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```

### **19.6.5 Combining bullet chart and sparklines**

Similarly, we can combining the bullet chart and sparklines using the steps below.

```{r}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 
```

```{r}
sales_data = sales_data %>%
  left_join(bullet)
```

```{r}
#| echo: true
#| message: false
#| warning: false
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## **19.7 Interactive Information Dashboard Design: reactable and reactablefmtr methods**

In this section, you’ll learn how to create **interactive dashboards** using the **`reactable`** and **`reactablefmtr`** packages.\
Before you begin, it is **strongly recommended** to visit their official webpages and go through the available materials.\
You **don’t need to memorize everything**—just get a general overview of their **features and purposes**.

To build **interactive sparklines**, you’ll also need to install the **`dataui`** package using the code chunk below:

```{r}
remotes::install_github("timelyportfolio/dataui")
```

Next, you all need to load the package onto R environment by using the code chunk below.

```{r}
library(dataui)
```

### **19.7.1 Plotting interactive sparklines**

Similar to gtExtras, to plot an interactive sparklines by using reactablefmtr package we need to prepare the list field by using the code chunk below.

```{r}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

Next, [react_sparkline](https://kcuilla.github.io/reactablefmtr/reference/react_sparkline.html) will be to plot the sparklines as shown below.

```{r}
#| echo: true
#| message: false
#| warning: false
reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

### **19.7.2 Changing the pagesize**

By default the pagesize is 10. In the code chunk below, arguments defaultPageSize is used to change the default setting.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

### **19.7.3 Adding points and labels**

In the code chunk below `highlight_points` argument is used to show the minimum and maximum values points and `label` argument is used to label first and last values.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```

### **19.7.4 Adding reference line**

In the code chunk below `statline` argument is used to show the mean line.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```

### **19.7.5 Adding bandline**

Instead adding reference line, bandline can be added by using the `bandline`argument.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```

### **19.7.6 Changing from sparkline to sparkbar**

Instead of displaying the values as sparklines, we can display them as sparkbars as shiwn below.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```
