---
title: "Hands-on_Ex17"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **17  Analytical Mapping**

## **17.1 Overview**

### **17.1.1 Objectives**

In this in-class exercise, you will gain hands-on experience on using appropriate R methods to plot analytical maps.

### **17.1.2 Learning outcome**

By the end of this in-class exercise, you will be able to use appropriate **tmap** and **tidyverse** functions to:

-   Import geospatial data in **rds** format into the R environment.

-   Create **cartographic-quality choropleth maps** using relevant **tmap** functions.

-   Generate a **rate map** to visualize proportional data.

-   Construct a **percentile map** for comparative analysis.

-   Develop a **boxmap** to identify spatial outliers.

## **17.2 Getting Started**

### **17.2.1 Installing and loading packages**

```{r}
pacman::p_load(tmap, tidyverse, sf)
```

### **17.2.2 Importing data**

For this hands-on exercise, we will use a pre-prepared dataset called **NGA_wp.rds**. This dataset is a polygon feature data frame containing information on water points in Nigeria at the **LGA (Local Government Area) level**. You can find the dataset in the **rds** subdirectory within the hands-on data folder.

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

## **17.3 Basic Choropleth Mapping**

### **17.3.1 Visualising distribution of non-functional water point**

```{r}
p1 <- tm_shape(NGA_wp) +
  tm_fill("wp_functional",
          n = 10,
          style = "equal",
          palette = "Blues") +
  tm_borders(lwd = 0.1,
             alpha = 1) +
  tm_layout(main.title = "Distribution of functional water point by LGAs",
            legend.outside = FALSE)
```

```{r}
p2 <- tm_shape(NGA_wp) +
  tm_fill("total_wp",
          fill.scale = tm_scale_intervals(style = "equal"),
          values = "brewer.blues") +
  tm_borders(lwd = 0.1, fill_alpha = 1) +
  tm_title("Distribution of total water point by LGAs") +
  tm_layout(legend.outside = FALSE)
```

```{r}
#| fig.width: 12
#| fig.height: 7
tmap_arrange(p2, p1, nrow = 1)
```

## **17.4 Choropleth Map for Rates**

Mapping rates instead of raw counts is crucial because water points are not evenly distributed across different areas. If we only map the total number of water points, we may misinterpret the data, as larger areas might naturally have more water points. Instead, using rates allows us to normalize the data, ensuring a more meaningful and accurate representation of water accessibility in different regions.

### **17.4.1 Deriving Proportion of Functional Water Points and Non-Functional Water Points**

We will tabulate the proportion of functional water points and the proportion of non-functional water points in each LGA. In the following code chunk, `mutate(`) from **dplyr** package is used to derive two fields, namely *pct_functional* and *pct_nonfunctional*.

```{r}
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

### **17.4.2 Plotting map of rate**

```{r}
#| echo: true
#| message: false
#| warning: false
tm_shape(NGA_wp) +
  tm_fill("pct_functional",
          legend.hist = TRUE,
          fill.scale = tm_scale_intervals(style = "equal", 
                                          values = "brewer.blues")
          ) +
  tm_borders(lwd = 0.1, fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs") +
  tm_layout(legend.outside = TRUE)


```

## **17.5 Extreme Value Maps**

Extreme value maps are a specialized type of choropleth map designed to emphasize outliers at both the low and high ends of a dataset. Their primary purpose is to highlight extreme values rather than display a smooth distribution. This approach is inspired by spatial exploratory data analysis (EDA), incorporating spatial features into traditional non-spatial EDA methods (Anselin, 1994).

### **17.5.1 Percentile Map**

The percentile map is a special type of quantile map with six specific categories: 0-1%,1-10%, 10-50%,50-90%,90-99%, and 99-100%. The corresponding breakpoints can be derived by means of the base R quantile command, passing an explicit vector of cumulative probabilities as c(0,.01,.1,.5,.9,.99,1). Note that the begin and endpoint need to be included.

#### 17.5.1.1 Data Preparation

Step 1: Exclude records with NA by using the code chunk below.

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()
```

Step 2: Creating customised classification and extracting values

```{r}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

::: callout-important
When variables are extracted from an sf data.frame, the geometry is extracted as well. For mapping and spatial manipulation, this is the expected behavior, but many base R functions cannot deal with the geometry. Specifically, the `quantile()` gives an error. As a result `st_set_geomtry(NULL)` is used to drop geomtry field.
:::

#### 17.5.1.2 Why writing functions?

Using functions instead of copy-pasting code provides three key advantages:

1.  Improved Readability – A well-named function makes your code easier to understand.

2.  Easier Maintenance – When requirements change, you only need to update the function in one place rather than modifying multiple sections of code.

3.  Reduced Errors – Avoids mistakes that occur when manually copying and pasting, such as forgetting to update variable names consistently.

Source: [Chapter 19: Functions](https://r4ds.had.co.nz/functions.html#functions) of **R for Data Science**.

#### 17.5.1.3 Creating the get.var function

Firstly, we will write an R function as shown below to extract a variable (i.e. *wp_nonfunctional*) as a vector out of an sf data.frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

#### 17.5.1.4 A percentile mapping function

```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```

#### 17.5.1.5 Test drive the percentile mapping function

```{r}
percentmap("total_wp", NGA_wp)

```

Note that this is just a bare bones implementation. Additional arguments such as the title, legend positioning just to name a few of them, could be passed to customise various features of the map.

### **17.5.2 Box map**

A box map is an enhanced version of a quartile map, designed to highlight outliers. It adds extra categories for extreme values:

-   If lower outliers exist, the first break starts at the minimum value, and the second break is the lower fence.

-   If there are no lower outliers, the first break starts at the lower fence, and the second break is the minimum value(meaning no observations exist between them).

This method helps identify extreme values while maintaining the structure of a quartile-based classification.

```{r}
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()
```

-   Displaying summary statistics on a choropleth map by using the basic principles of boxplot.

-   To create a box map, a custom breaks specification will be used. However, there is a complication. The break points for the box map vary depending on whether lower or upper outliers are present.

#### 17.5.2.1 Creating the boxbreaks function

The code chunk below is an R function that creating break points for a box map.

-   arguments:

    -   v: vector with observations

    -   mult: multiplier for IQR (default 1.5)

-   returns:

    -   bb: vector with 7 break points compute quartile and fences

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

#### 17.5.2.2 Creating the get.var function

The code chunk below is an R function to extract a variable as a vector out of an sf data frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

#### 17.5.2.3 Test drive the newly created function

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

#### 17.5.2.4 Boxmap function

The following R function generates a **box map**, which highlights outliers based on quartiles and interquartile range (IQR).

**Function Arguments:**

-   **`vnam`** → Variable name (as a character, in quotes)

-   **`df`** → A **simple features polygon layer** (spatial data)

-   **`legtitle`** → Title for the **legend**

-   **`mtitle`** → Title for the **map**

-   **`mult`** → **Multiplier for IQR** to define fences for outliers

**Function Output:**

-   Returns a **tmap element** that plots a **box map**, visually distinguishing outliers from other data points.

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
```

```{r}
#| fig.width: 12
#| fig.height: 7
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```
