---
title: "Untitled"
---

```{r}
#| warning: false
pacman::p_load(tidyverse,tsibble,feasts,fable,seasonal)
```

### **Importing the data**

```{r}
ts_data <- read_csv("visitor_arrivals_by_air.csv",show_col_types = FALSE)
```

`dmy()` of **lubridate** package 

```{r}
ts_data$`Month-Year` <- dmy(ts_data$`Month-Year`)
```

```{r}
ts_data
```

```{r}
ts_data_ts <- ts(ts_data)       
head(ts_data_ts)
```

### **Converting `tibble` object to `tsibble`object**

```{r}
ts_tsibble <- ts_data %>%
  mutate(Month = yearmonth(`Month-Year`)) %>%
  as_tsibble(index = `Month`)
```

## **Visualising Time-series Data**

```{r}
ts_longer <- ts_data %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

### **Visualising single time-series: ggplot2 methods**

```{r}
ts_longer %>%
  filter(Country == "Vietnam") %>%
  ggplot(aes(x = `Month-Year`, 
             y = Arrivals))+
  geom_line(size = 0.5)
```

### **Plotting multiple time-series data with ggplot2 methods**

```{r}
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals,
           color = Country))+
  geom_line(size = 0.5) +
  theme(legend.position = "bottom", 
        legend.box.spacing = unit(0.5, "cm"))
```

In order to provide effective comparison, [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) of **ggplot2**package is used to create small multiple line graph also known as trellis plot.

```{r}
#| echo: false
#| fig-height: 23
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals))+
  geom_line(size = 0.5) +
  facet_wrap(~ Country,
             ncol = 3,
             scales = "free_y") +
  theme_bw()
```

## **Visual Analysis of Time-series Data**

```{r}
tsibble_longer <- ts_tsibble %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

### **Visual Analysis of Seasonality with Seasonal Plot**

```{r}
tsibble_longer %>%
  filter(Country == "Italy" |
         Country == "Vietnam" |
         Country == "United Kingdom" |
         Country == "Germany") %>% 
  gg_season(Arrivals)
```

### **Visual Analysis of Seasonality with Cycle Plot**

```{r}
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  autoplot(Arrivals) + 
  facet_grid(Country ~ ., scales = "free_y")
```

```{r}
#| fig-height: 6
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  gg_subseries(Arrivals)
```

### **Multiple time-series decomposition**

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  PACF(Arrivals) %>%
  autoplot()
```

### **Visual STL diagnostics with feasts**

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```

### **Classical Decomposition with feasts**

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  model(
    classical_decomposition(
      Arrivals, type = "additive")) %>%
  components() %>%
  autoplot()
```

### **Time Series Data Sampling**

```{r}
vietnam_ts <- tsibble_longer %>%
  filter(Country == "Vietnam") %>% 
  mutate(Type = if_else(
    `Month-Year` >= "2019-01-01", 
    "Hold-out", "Training"))

vietnam_train <- vietnam_ts %>%
  filter(`Month-Year` < "2019-01-01)
```

### **`Exploratory Data Analysis (EDA): Time Series Data`**

```{r}
vietnam_train %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```

### **`Fitting forecasting models-code change`**

```{r}
ETS(y ~ error("A") + trend("N") + season("A"))


```

```{r}
fit_ses <- vietnam_train %>%
  model(ETS(Arrivals ~ error("A") 
            + trend("N") 
            + season("N")))
fit_ses
```

```{r}
gg_tsresiduals(fit_ses)
```

```{r}
fit_ses %>%
  report()
```

```{r}
Vietnam_WH <- vietnam_train %>%
  model(
    Additive = ETS(Arrivals ~ error("A") 
                   + trend("A") 
                   + season("A")),
    Multiplicative = ETS(Arrivals ~ error("M") 
                         + trend("A") 
                         + season("M"))
    )

Vietnam_WH %>% report()
```

```{r}
fit_ETS <- vietnam_train %>%
  model(`SES` = ETS(Arrivals ~ error("A") + 
                      trend("N") + 
                      season("N")),
        `Holt`= ETS(Arrivals ~ error("A") +
                      trend("A") +
                      season("N")),
        `damped Holt` = 
          ETS(Arrivals ~ error("A") +
                trend("Ad") + 
                season("N")),
        `WH_A` = ETS(
          Arrivals ~ error("A") + 
            trend("A") + 
            season("A")),
        `WH_M` = ETS(Arrivals ~ error("M") 
                         + trend("A") 
                         + season("M"))
  )
```

```{r}
fit_ETS %>%
  forecast(h = "12 months") %>%
  autoplot(vietnam_ts, 
           level = NULL)
```

```{r}
fit_autoETS <- vietnam_train %>%
  model(ETS(Arrivals))
fit_autoETS %>% report()
```

```{r}
gg_tsresiduals(fit_autoETS)

```

```{r}
fit_autoETS %>%
  forecast(h = "12 months") %>%
  autoplot(vietnam_train)
```

## **19.8 AutoRegressive Integrated Moving Average(ARIMA) Methods for Time Series Forecasting: fable (tidyverts) methods**

```{r}
vietnam_train %>%
  gg_tsdisplay(plot_type='partial')
```

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  ACF(Arrivals) %>% 
  autoplot()
```

```{r}
tsibble_longer %>%
  filter(`Country` == "United Kingdom") %>%
  ACF(Arrivals) %>% 
  autoplot()
```

```{r}
tsibble_longer %>%
  filter(Country == "Vietnam") %>%
  gg_tsdisplay(difference(
    Arrivals,
    lag = 1), 
    plot_type='partial')
```

```{r}
tsibble_longer %>%
  filter(Country == "Vietnam") %>%
  gg_tsdisplay(difference(
    Arrivals,
    difference = 12), 
    plot_type='partial')
```

```{r}
fit_arima <- vietnam_train %>%
  model(
    arima200 = ARIMA(Arrivals ~ pdq(2,0,0)),
    sarima210 = ARIMA(Arrivals ~ pdq(2,0,0) + 
                        PDQ(2,1,0))
    )
report(fit_arima)
```

```{r}
fit_autoARIMA <- vietnam_train %>%
  model(ARIMA(Arrivals))
report(fit_autoARIMA)
```

```{r}
bind_rows(
    fit_autoARIMA %>% accuracy(),
    fit_autoETS %>% accuracy(),
    fit_autoARIMA %>% 
      forecast(h = 12) %>% 
      accuracy(vietnam_ts),
    fit_autoETS %>% 
      forecast(h = 12) %>% 
      accuracy(vietnam_ts)) %>%
  select(-ME, -MPE, -ACF1)
```

```{r}
ASEAN <- tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Malaysia" |
         Country == "Indonesia" |
         Country == "Thailand" |
         Country == "Philippines")
```

```{r}
ASEAN_train <- ASEAN %>%
  mutate(Type = if_else(
    `Month-Year` >= "2019-01-01", 
    "Hold-out", "Training")) %>%
  filter(Type == "Training")
```

```{r}
ASEAN_fit <- ASEAN_train %>%
  model(
    ets = ETS(Arrivals),
    arima = ARIMA(Arrivals)
  )
```

```{r}
ASEAN_fit %>%
  glance()
```

### **19.8.10 Extracintg fitted and residual values**

```{r}
ASEAN_fit %>%
  augment()
```

```{r}
ASEAN_fit %>%
  accuracy() %>%
  arrange(Country)
```

```{r}
ASEAN_fc <- ASEAN_fit %>%
  forecast(h = "12 months")
```

```{r}
ASEAN_fc %>%
  autoplot(ASEAN)
```
