---
title: "Untitled"
---

```{r}
#| warning: false
pacman::p_load(tidyverse,tsibble,feasts,fable,seasonal)
```

### **Importing the data**

```{r}
ts_data <- read_csv("visitor_arrivals_by_air.csv",show_col_types = FALSE)
```

`dmy()` of **lubridate** package 

```{r}
ts_data$`Month-Year` <- dmy(ts_data$`Month-Year`)
```

```{r}
ts_data
```

```{r}
ts_data_ts <- ts(ts_data)       
head(ts_data_ts)
```

### **Converting `tibble` object to `tsibble`object**

```{r}
ts_tsibble <- ts_data %>%
  mutate(Month = yearmonth(`Month-Year`)) %>%
  as_tsibble(index = `Month`)
```

##  **Visualising Time-series Data**

```{r}
ts_longer <- ts_data %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

### **Visualising single time-series: ggplot2 methods**

```{r}
ts_longer %>%
  filter(Country == "Vietnam") %>%
  ggplot(aes(x = `Month-Year`, 
             y = Arrivals))+
  geom_line(size = 0.5)
```

### **Plotting multiple time-series data with ggplot2 methods**

```{r}
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals,
           color = Country))+
  geom_line(size = 0.5) +
  theme(legend.position = "bottom", 
        legend.box.spacing = unit(0.5, "cm"))
```

In order to provide effective comparison, [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) of **ggplot2**package is used to create small multiple line graph also known as trellis plot.

```{r}
#| echo: false
#| fig-height: 12
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals))+
  geom_line(size = 0.5) +
  facet_wrap(~ Country,
             ncol = 3,
             scales = "free_y") +
  theme_bw()
```
