---
title: "Hands-on_Ex14"
author: "Jeounghyun Kim"
date-modified: "last-modified" 
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

# **14  Visualising and Analysing Time-oriented Data**

## **14.1 Learning Outcome**

By the end of this hands-on exercise you will be able create the followings data visualisation by using R packages:

-   plotting a calender heatmap by using ggplot2 functions,

-   plotting a cycle plot by using ggplot2 function,

-   plotting a slopegraph

-   plotting a horizon chart

## **14.2 Getting Started**

Write a code chunk to check, install and launch the following R packages: scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table and tidyverse.

```{r}
#| message: false
#| warning: false
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table, CGPfunctions, ggHoriPlot, tidyverse)
```

## **14.3 Plotting Calendar Heatmap**

#### **Learning Outcomes**

By the end of this section, you will be able to:

-   Plot a calendar heatmap using `ggplot2` and its extensions.

-   Write functions in R to automate tasks.

-   Extract and manipulate date/time-related fields using base R and the `lubridate` package.

-   Perform data preparation using the `tidyr` and `dplyr` packages.

### **14.3.1 The Data**

For the purpose of this hands-on exercise, *eventlog.csv* file will be used. This data file consists of 199,999 rows of time-series cyber attack records by country.

### **14.3.2 Importing the data**

First, you will use the code chunk below to import *eventlog.csv* file into R environment and called the data frame as *attacks*.

```{r}
attacks <- read_csv("data/eventlog.csv")
```

### **14.3.3 Examining the data structure**

Let's examine the imported data frame before proceeding with further analysis. The `kable()` function can be used to review its structure.

```{r}
#| echo: true
#| message: false
#| warning: false
kable(head(attacks))
```

The data frame consists of three columns: `timestamp`, `source_country`, and `tz`.

-   The `timestamp` column stores date-time values in POSIXct format.

-   The `source_country` column represents the source of the attack using ISO 3166-1 alpha-2 country codes.

-   The `tz` column indicates the time zone of the source IP address.

### **14.3.4 Data Preparation**

Step 1: Deriving Weekday and Hour of the Day Fields

Before plotting the calendar heatmap, we need to create two new fields:

-   **`wkday`**: Represents the day of the week.

-   **`hour`**: Represents the hour of the day.

To achieve this, we will write a function to extract these fields using the following functions:

-   **`ymd_hms()`** and **`hour()`** from the **`lubridate`** package to handle date-time conversion and extract the hour.

-   **`weekdays()`** from **base R** to retrieve the weekday from a date.

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts, 
                        tz = tz[1], 
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
  }
```

Step 2: Deriving the attacks tibble data frame

```{r}
#| message: false
#| warning: false
wkday_levels <- c('Saturday', 'Friday', 
                  'Thursday', 'Wednesday', 
                  'Tuesday', 'Monday', 
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp, 
                   .$source_country, 
                   .$tz)) %>% 
  ungroup() %>% 
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour  = factor(
      hour, levels = 0:23))
```

::: callout-note
Beside extracting the necessary data into *attacks* data frame, `mutate()` of **dplyr** package is used to convert *wkday* and *hour* fields into **factor** so they’ll be ordered when plotting
:::

Shows the tidy tibble table after processing.

```{r}
kable(head(attacks))
```

### **14.3.5 Building the Calendar Heatmaps**

::: panel-tabset
#### The Plot

```{r}
#| echo: false
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```

#### The code chunk

```{r, echo=TRUE, eval=FALSE}
grouped <- attacks %>% 
  count(wkday, hour) %>% 
  ungroup() %>%
  na.omit()

ggplot(grouped, 
       aes(hour, 
           wkday, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() +
scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = NULL, 
     y = NULL, 
     title = "Attacks by weekday and time of day") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 6) )
```
:::

#### Things to learn from the code chunk

-   **Data Aggregation**

    -   The `grouped` tibble is created by aggregating data based on `wkday` (weekday) and `hour` fields.

    -   A new field `n` is derived using `group_by()` and `count()` functions.

-   **Handling Missing Values**

    -   `na.omit()` is used to remove missing values from the dataset.

-   **Plotting Tiles (Grids)**

    -   `geom_tile()` is used to create a heatmap by placing tiles at each `x` and `y` position.

    -   The arguments `color` and `size` define the border color and line thickness of the tiles.

-   **Theme Customization**

    -   `theme_tufte()` from the **ggthemes** package is used to remove unnecessary visual elements from the default ggplot2 chart.

    -   To compare with the default theme, you can **comment out this line** and observe the changes.

-   **Aspect Ratio Adjustment**

    -   `coord_equal()` ensures the plot maintains a **1:1 aspect ratio**, preventing distortion.

-   **Color Gradient for Visualization**

    -   `scale_fill_gradient()` applies a **two-color gradient** (from low to high values) to enhance data readability.

Since we have values for every combination of `hour` and `wkday`, we can directly group the data by these fields and plot it without additional preprocessing. This simplifies the workflow, allowing us to proceed with visualization immediately.

### **14.3.6 Building Multiple Calendar Heatmaps**

**Challenge:** Building multiple heatmaps for the top four countries with the highest number of attacks.

```{r}
#| echo: false
#1
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
top4 <- attacks_by_country$source_country[1:4]
#2
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
#3
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

### **14.3.7 Plotting Multiple Calendar Heatmaps**

Step 1: Deriving Attack by Country Object

To identify the top 4 countries with the highest number of attacks, follow these steps:

1.  Count the number of attacks for each country.

2.  Calculate the percentage of attacks contributed by each country.

3.  Store the results in a tibble data frame for further analysis.

```{r}
#| include: false
attacks_by_country <- count(
  attacks, source_country) %>%
  mutate(percent = percent(n/sum(n))) %>%
  arrange(desc(n))
```

Step 2: Preparing the tidy data frame

In this step, you are required to extract the attack records of the top 4 countries from *attacks*data frame and save the data in a new tibble data frame (i.e. *top4_attacks*).

```{r}
#| message: false
#| warning: false
#| include: false
top4 <- attacks_by_country$source_country[1:4]
top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

### **14.3.8 Plotting Multiple Calendar Heatmaps**

Step 3: Plotting the Multiple Calender Heatmap by using ggplot2 package.

```{r}
#| echo: true
#| message: false
#| warning: false
ggplot(top4_attacks, 
       aes(hour, 
           wkday, 
           fill = n)) + 
  geom_tile(color = "white", 
          size = 0.1) + 
  theme_tufte(base_family = "Helvetica") + 
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                    low = "sky blue", 
                    high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL, 
     title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6) )
```

## **14.4 Plotting Cycle Plot**

In this section, you will learn how to create a **cycle plot** to visualize time-series patterns and trends in **visitor arrivals from Vietnam** using **ggplot2** in R.

By the end of this section, you will be able to:

-   Use **ggplot2** functions to create a cycle plot.

-   Identify seasonal patterns and trends in time-series data.

-   Effectively communicate insights using visualizations.

```{r}
# str(Vietnam$month)  

```

```{r}
#| echo: false
#| message: false
#| warning: false
#1
air <- read_excel("data/arrivals_by_air.xlsx")

air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))

Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)


hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))

ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")

```

### **14.4.1 Step 1: Data Import**

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

### **14.4.2 Step 2: Deriving month and year fields**

Next, two new fields called *month* and *year* are derived from *Month-Year* field.

```{r}
air$month <- factor(month(air$`Month-Year`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 
air$year <- year(ymd(air$`Month-Year`))
```

### **14.4.3 Step 4: Extracting the target country**

Next, extract data for the target country (i.e. Vietnam)

```{r}
Vietnam <- air %>% 
  select(`Vietnam`, 
         month, 
         year) %>%
  filter(year >= 2010)
```

### **14.4.4 Step 5: Computing year average arrivals by month**

Using **`group_by()`** and **`summarise()`** from dplyr, we calculate the average monthly visitor arrivals across years to analyze seasonal trends.

```{r}
hline.data <- Vietnam %>% 
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### **14.4.5 Step 6: Plotting the cycle plot**

```{r}
#| message: false
#| warning: false
ggplot() + 
  geom_line(data=Vietnam,
            aes(x=year, 
                y=`Vietnam`, 
                group=month), 
            colour="black") +
  geom_hline(aes(yintercept=avgvalue), 
             data=hline.data, 
             linetype=6, 
             colour="red", 
             size=0.5) + 
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```

## **14.5 Plotting Slopegraph**

In this section, you will learn how to create a **slopegraph** in R.

**Before starting:**

-   Ensure the **CGPfunctions** package is installed and loaded.

-   Refer to **`newggslopegraph()`** documentation for usage and arguments.

### **14.5.1 Step 1: Data Import**

```{r}
rice <- read_csv("data/rice.csv")
```

### **14.5.2 Step 2: Plotting the slopegraph**

```{r}
#| echo: true
#| message: false
#| warning: false
rice %>% 
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  newggslopegraph(Year, Yield, Country,
                Title = "Rice Yield of Top 11 Asian Counties",
                SubTitle = "1961-1980",
                Caption = "Prepared by: Dr. Kam Tin Seong")
```

::: callout-tip
#### Thing to learn from the code chunk above

For effective data visualisation design, `factor()` is used convert the value type of *Year* field from numeric to factor.
:::
